The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multi-Phase Forest Inventories

### 2017-10-12 JStatSoft_forestinventory_master-concordance.tex
readyforsubmission

\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
1 108 1 1 4 1 1 1 4 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 116:%
1 98 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 215:%
124 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 220:%
1 68 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 289:%
130 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 294:%
1 109 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 7 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 126 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 810:%
136}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 811:%
1 105 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 151 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1231:%
138 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1236:%
1 35 1 1 2 11 0 1 2 3 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1288:%
144 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1293:%
1 12 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 11 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 13 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 18 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1542:%
150 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1547:%
1 9 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 9 1 1 4 1 1 1 2 11 0 1 2 16 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1742:%
156 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1747:%
1 11 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1759:%
162 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1764:%
1 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1769:%
168 11 1}

>>>>

\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
1 108 1 1 4 1 1 1 4 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 116:%
1 98 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 215:%
124 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 220:%
1 67 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 288:%
130 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 293:%
1 104 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 7 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 139 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 817:%
136}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 818:%
1 105 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 150 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1237:%
138 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1242:%
1 35 1 1 2 11 0 1 2 3 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1294:%
144 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1299:%
1 12 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 11 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 13 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 18 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1548:%
150 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1553:%
1 9 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 9 1 1 4 1 1 1 2 11 0 1 2 16 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1748:%
156 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1753:%
1 11 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1765:%
162 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1770:%
1 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1775:%
168 11 1}

### 2017-10-12 rnw/Future Plans.Rnw
readyforsubmission

The \pkg{forestinventory} package currently provides a fairly well-rounded toolkit for forestry inventorists to integrate auxiliary information into their estimates using the model-assisted methods under the design-based approach.  Although 32 combinations of inventory scenarios, estimators and sample designs are covered, there are still potential improvements planned for the future. As this is an open-source project, everyone is encouraged to give feedback and/or make contributions on the package's development page on GitHub. Currently planned extensions include:

>>>>

The \pkg{forestinventory} package currently provides a fairly well-rounded toolkit for forestry inventorists to integrate auxiliary information into their estimates using the model-assisted methods under the design-based approach.  Although 32 combinations of inventory scenarios, estimators and sample designs are covered, there are still potential improvements planned for the future. As this is an open-source project, everyone is encouraged to give feedback and/or make contributions on the package's development page on GitHub \citep{github_forestinventory}. Currently planned extensions include:

### 2017-10-12 rnw/Main structure and modules of the Package.Rnw
update Andreas 12.10.2017(2)
update Andreas 12.10.2017(3)
readyforsubmission

Most forest inventories gather terrestrial information by sending field crews to randomly (or systematically) selected points in the forest area defined by GPS coordinates. The crew then defines a sample plot by measuring individual trees located within one or multiple constructed inclusion circles around the sample point $x$, and aggregating their characteristics (e.g. timber volumes) to local plot densities (e.g. the timber density in $m^3/ha$). The estimators implemented in \pkg{forestinventory} use the so called \textit{infinite population approach} in order to bridge this inventory process to the mathematics behind the estimators. This approach assumes that the spatial distribution of the local density, $Y(x)$, over the forest area is determined by a fixed piecewise constant function, as visualized in Fig. \ref{fig:inf_pop_apr}. The \textit{population total} of the target variable (e.g. the total timber volume of the forest) is mathematically equivalent to the integral of that density function, which is depicted in Fig. \ref{fig:inf_pop_apr} as the volume under the density surface. From this perspective, the practical challenge is that the form of this function is unknown. Theoretically, we could get the total timber volume by observing the function value, i.e. the local density $Y(x)$, at each possible point $x$ over the forest area and taking their sum. However, this is impossible because there is an infinite number of points in the forest area. Our strategy is thus to take a sample of points, $s_2$, from an infinite population of possible points and use their associated local densities, i.e. $Y(x)$, to \textit{estimate} the integral $Y=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$ with $\hat{Y}=\frac{1}{n_2}\sum_{x \in s_2}Y(x)$.  The total timber volume can then be obtained by multiplying $\hat{Y}$ by the surface area of the forest, $\lambda(F)$.  All estimators included in \pkg{forestinventory} rest upon this theoretical perspective. The key point in the \textit{infinite population approach} is that a local density value $Y(x)$ is associated with the sample point $x$, which constitutes the sample unit, and not with the sample plot area. This has some theoretical advantages over the \textit{finite population approach}, where the sample units are the actual plot areas usually assumed to be either circular or rectangular.  This is mainly due to the impossibility of a perfect tesselation over an amorphous forest area by any choice of plot shape. Hence, the population in the finite approach is, strictly speaking, not well defined with respect to the forest area. The consideration of an underlying infinite population of sample points will also play an important role when incorporating auxiliary information in the frame of two- and three-phase estimation methods.

@@

\begin{figure}[htb]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
    \resizebox{1\hsize}{!}{\includegraphics{fig/phases_graphic(new).eps}}%
		\caption{} \label{fig:concmphase_and_sae_a}
		\end{subfigure}
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
	 \resizebox{0.58\hsize}{!}{\includegraphics{fig/sae_graphic.PNG}}%
   %\resizebox{0.6\hsize}{!}{\includegraphics{fig/sae_graphic(new).eps}}%
		\caption{} \label{fig:concmphase_and_sae_b}
	\end{subfigure}
\caption{\textbf{(a)} Concept of multiphase sampling. The square represents the forest area for which an inventory is being conducted. The points denote the sample locations $x$. Filled points indicate \textit{available} information. \textbf{(b)} Illustration of the Small Area Estimation problem}
\label{fig:concmphase_and_sae}
\end{figure}

@@

In the \pkg{forestinventory} package, estimators for two-phase and three-phase sampling are applied with the \code{twophase()} and \code{threephase()} functions. From these two overall function calls, various estimators for specific inventory scenarios under the chosen sampling design can be applied (Fig. \ref{fig:struct_package}). The decision what estimator to use thereby follows a tree-like structure which can serve the user as a guideline throughout this article as well as in future applications. The basic decision to make is whether an estimate and its variance should be computed for an entire inventory area (\textit{global estimators}) or only for subregions of the entire inventory area (\textit{small area estimators}). In the second case, the package offers three small area estimators that will in detail be described in the following sections. The estimators are available under \textit{exhaustive} and \textit{non-exhaustive} use of the auxiliary data. Additionally, the package also can calculate one-phase estimates solely based on terrestrial samples. All estimators are also available for \textit{cluster} sampling, in which case a sample unit consists of multiple, spatially agglomerated samples. The following sections describe the mathematical details and the application of the multi-phase estimators implemented in the \proglang{R} package \pkg{forestinventory}. While \cite{mandallaz2008, mandallaz2013techa, mandallaz2013techb,mandallaz2015tech} provides an extensive derivation of all estimators, we will provide the mathematical formulas that are actually implemented in the package. We will also restrict discussion to simple sampling. A special case under cluster sampling is described in Section \ref{sec:speccas_and_scen}.

\begin{figure}[htb]
\centering
\resizebox{1\hsize}{!}{\includegraphics{fig/Package_structure_4.PNG}}
\caption{Structure of the multi-phase estimators in the \proglang{R} package \pkg{forestinventory}}
\label{fig:struct_package}
\end{figure}

>>>>

Most forest inventories gather terrestrial information by sending field crews to randomly (or systematically) selected points in the forest area defined by GPS coordinates. The crew then defines a sample plot by measuring individual trees located within one or multiple constructed inclusion circles around the sample point $x$, and aggregating their characteristics (e.g. timber volumes) to local plot densities (e.g. the timber density in $m^3/ha$). The estimators implemented in \pkg{forestinventory} use the so called \textit{infinite population approach} in order to bridge this inventory process to the mathematics behind the estimators. This approach assumes that the spatial distribution of the local density, $Y(x)$, over the forest area is determined by a fixed piecewise constant function, as visualized in Fig. \ref{fig:inf_pop_apr}. The \textit{population total} of the target variable (e.g. the total timber volume of the forest) is mathematically equivalent to the integral of that density function, which is depicted in Fig. \ref{fig:inf_pop_apr} as the volume under the density surface. From this perspective, the practical challenge is that the form of this function is unknown. Theoretically, we could get the total timber volume by observing the function value, i.e. the local density $Y(x)$, at each possible point $x$ over the forest area and taking their sum. However, this is impossible because there is an infinite number of points in the forest area. Our strategy is thus to take a sample of points, $s_2$, from an infinite population of possible points and use their associated local densities, i.e. $Y(x)$, to \textit{estimate} the integral $Y=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$ with $\hat{Y}=\frac{1}{n_2}\sum_{x \in s_2}Y(x)$.  The total timber volume can then be obtained by multiplying $\hat{Y}$ by the surface area of the forest, $\lambda(F)$.  All estimators included in \pkg{forestinventory} rest upon this theoretical perspective. The key point in the \textit{infinite population approach} is that a local density value $Y(x)$ is associated with the sample point $x$, which constitutes the sample unit, and not with the sample plot area. This has some theoretical advantages over the \textit{finite population approach}, where the sample units are the actual plot areas usually assumed to be either circular or rectangular.  This is mainly due to the impossibility of a perfect tessellation over an amorphous forest area by any choice of plot shape. Hence, the population in the finite approach is, strictly speaking, not well defined with respect to the forest area. The consideration of an underlying infinite population of sample points will also play an important role when incorporating auxiliary information in the frame of two- and three-phase estimation methods.

@@

\begin{figure}[htb]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
    \resizebox{1\hsize}{!}{\includegraphics{fig/phases_graphic(new).eps}}%
		\caption{} \label{fig:concmphase_and_sae_a}
		\end{subfigure}
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
	 \resizebox{0.58\hsize}{!}{\includegraphics{fig/sae_graphic.PNG}}%
		\caption{} \label{fig:concmphase_and_sae_b}
	\end{subfigure}
\caption{\textbf{(a)} Concept of multiphase sampling. The square represents the forest area for which an inventory is being conducted. The points denote the sample locations $x$. Filled points indicate \textit{available} information. \textbf{(b)} Illustration of the Small Area Estimation problem}
\label{fig:concmphase_and_sae}
\end{figure}

@@

In the \pkg{forestinventory} package, estimators for two-phase and three-phase sampling are applied with the \code{twophase()} and \code{threephase()} functions. From these two overall function calls, various estimators for specific inventory scenarios under the chosen sampling design can be applied (Fig. \ref{fig:struct_package}). Choosing an estimator follows a tree-like structure which can serve the user as a guideline throughout this article as well as in future applications. The basic decision to make is whether an estimate and its variance should be computed for an entire inventory area (\textit{global estimators}) or only for subregions of the entire inventory area (\textit{small area estimators}). In the second case, the package offers three small area estimators that will in detail be described in the following sections. The estimators are available under \textit{exhaustive} and \textit{non-exhaustive} use of the auxiliary data. Additionally, the package also can calculate one-phase estimates solely based on terrestrial samples. All estimators are also available for \textit{cluster} sampling, in which case a sample unit consists of multiple, spatially agglomerated samples. The following sections describe the mathematical details and the application of the multi-phase estimators implemented in the \proglang{R} package \pkg{forestinventory}. While \cite{mandallaz2008, mandallaz2013techa, mandallaz2013techb,mandallaz2015tech} provides an extensive derivation of all estimators, we will provide the mathematical formulas that are actually implemented in the package. We will also restrict discussion to simple sampling, while the formulas for cluster sampling are available in the technical reports \citep{mandallaz2016, mandallaz2013techa, mandallaz2013techb}. A special case under cluster sampling is described in Section \ref{sec:speccas_and_scen}.

\begin{figure}[htb]
\centering
\resizebox{1\hsize}{!}{\includegraphics{fig/Package_structure_5.eps}}
\caption{Structure of the multi-phase estimators in the \proglang{R} package \pkg{forestinventory}}
\label{fig:struct_package}
\end{figure}

### 2017-10-12 rnw/Special Cases and Scenarios.Rnw
update Andreas 12.10.2017(3)
readyforsubmission

To demonstrate \textit{post-stratification}, we first create an artificial categorical variable \textit{developement stage} (\code{stage}) by clustering the mean canopy heights of the \code{grisons} data set into 3 height classes:

@@

\textit{Two-phase sampling for stratification} is applied if the model only contains categorical variables, in this case the factor variable \code{stage}. Linear regression models only fitted with categorical variables produce ANOVA models, which when used in multi-phase regression estimators is equivalent to post-stratification. For our example, this means that the model predictions are simply the means of the terrestrial response values within each development stage (within-strata means).

@@

As mentioned in Section \ref{sec:packstruc}, \textit{cluster sampling} is a special case of sample designs where the sample consists of more than one spatially agglomerated sample points. One randomly places the sample location $x$ in the inventory area as in the simple sampling design, but then $M-1$ additional sample locations $x_2,...,x_M$ are created close to the cluster origin $x$ by adding a fixed set of spatial vectors $e_2,...,e_M$ to $x$. The idea of cluster sampling is to increase the amount of information without increasing the travelling costs of the terrestrial campaign. However, the information gathered at all sub-locations of a cluster is then averaged on the cluster level, and this average value then references exactly one point, i.e. the cluster origin $x$. Without going into too much mathematical detail, the estimators under simple sampling are thus extended in a way that all parameters (local density, mean vector of explanatory variables, mean model residuals) have to be calculated as the weighted cluster means with $M(x)$ being the cluster weights. Whereas the geometric form and the number of sample locations per cluster $M$ is fixed (i.e. defined by the inventorist), the actual number of points $M(x)$ falling into the forest area $F$ at sample location $x$ is random because the cluster origin $x$ is random. The \pkg{forstinventory} package identifies clusters via a unique cluster ID that is assigned to a column in the input data set. Its column name is passed to the argument \code{cluster} in the \code{twophase()} and \code{threephase()} function calls.

@@

However, if an $s_2$ and/or $s_1$ point misses a variable which is only used in the full regression model (in this example \code{q75}), the function will \textbf{recode} the phase indicator of that point to $s_0$, since the point still provides the required information for the reduced model. If this concerns an $s_2$ sample location, the associated response variable can no longer be used.

@@

If a categorical variable is used in the regression model(s) and the terrestrial sample $s_2$ is considerably small compared to the $s_1$ phase, it might occur that a category is only present in the $s_1 \backslash s_2$ sample, and thus missing in the $s_2$ sample. Conclusively, an internal regression model cannot be calculated and the function \textbf{stops} with the following error message:

>>>>

To demonstrate \textit{post-stratification}, we first create an artificial categorical variable \textit{development stage} (\code{stage}) by clustering the mean canopy heights of the \code{grisons} data set into 3 height classes:

@@

\textit{Two-phase sampling for stratification} is applied if the model only contains categorical variables, in this case the factor variable \code{stage}. Linear regression models only fitted with categorical variables produce ANOVA models, which when used in multi-phase regression estimators, is equivalent to post-stratification. For our example, this means that the model predictions are simply the means of the terrestrial response values within each development stage (within-strata means).

@@

As mentioned in Section \ref{sec:packstruc}, \textit{cluster sampling} is a special case of sample designs where the sample consists of more than one spatially agglomerated sample points. One randomly places the sample location $x$ in the inventory area as in the simple sampling design, but then $M-1$ additional sample locations $x_2,...,x_M$ are created close to the cluster origin $x$ by adding a fixed set of spatial vectors $e_2,...,e_M$ to $x$. The idea of cluster sampling is to increase the amount of information without increasing the travel costs of the terrestrial campaign. However, the information gathered at all sub-locations of a cluster is then averaged on the cluster level, and this average value then references exactly one point, i.e. the cluster origin $x$. Without going into too much mathematical detail, the estimators under simple sampling are thus extended in a way that all parameters (local density, mean vector of explanatory variables, mean model residuals) have to be calculated as the weighted cluster means with $M(x)$ being the cluster weights. Whereas the geometric form and the number of sample locations per cluster $M$ is fixed (i.e. defined by the inventorist), the actual number of points $M(x)$ falling into the forest area $F$ at sample location $x$ is random because the cluster origin $x$ is random. The \pkg{forstinventory} package identifies clusters via a unique cluster ID that is assigned to a column in the input data set. Its column name is passed to the argument \code{cluster} in the \code{twophase()} and \code{threephase()} function calls.

@@

However, if an $s_2$ and/or $s_1$ point is missing a variable which is only used in the full regression model (in this example \code{q75}), the function will \textbf{recode} the phase indicator of that point to $s_0$, since the point still provides the required information for the reduced model. If this concerns an $s_2$ sample location, the associated value of the response variable can no longer be used.

@@

If a categorical variable is used in the regression model(s) and the terrestrial sample $s_2$ is considerably small compared to the $s_1$ phase, it might occur that a category is only present in the $s_1 \backslash s_2$ sample, and thus missing in the $s_2$ sample. Therefore, an internal regression model cannot be calculated and the function \textbf{stops} with the following error message:
